### 2020-05-19 manuscript/manuscript.rmd
reworked intro

Considerable debate remains regarding the deepest split in the animal tree of life. This question has major implications for our understanding of the earliest events in animal evolution, including the origin of the nervous system. Some previous analyses found support for the first split being between comb jellies and all other animals (the Ctenophora-sister hypothesis), while others found support the first split being between sponges and all other animals (Porifera-sister). These studies differed in a many respects, including the genes and species selected, molecular evolution models, and analysis software. Here we systematically explore this question under consistent conditions by synthesizing data and results from XX previous phylogenomic analyses and performing new standardized analyses. Our reanalyses are consistent with previous studies that recover Porifera-sister with reduced outgroup (non-animal) sampling and complex site heterogeneous models, and Ctenophora-sister under other conditions. Though this had been interpreted by some to suggest Ctenophora-sister is an artifact of amino-acid composition bias, our analyses instead indicate that many site heterogeneous models do support Ctenophora-sister, and Porifera-sister is a result of unconstrained overfit models. Our analyses help reconcile variation across previous studies, and the datasets and analyses consolidated here will be a useful test-platform for phylogenetic methods development.

@@

Historically, there was little debate about the root of the animal tree of life and Porifera-sister (Fig. 1) was widely accepted though rarely tested. By contrast, there has long been uncertainty about the relationship of Ctenophora to other animals [@Wallberg:2004ws]. The first phylogenomic study to include ctenophores [@Dunn:2008ky] suggested a new hypothesis, now referred to as Ctenophora-sister, that ctenophores are our most distant animal relative. Since then many more studies have been published, some supporting Ctenophora-sister, some Porifera-sister, and some neither (Table 1). As the problem has become better characterized, it has become an interesting test-case to phylogenetic biologists beyond those concerned with this particular biological problem. Work has been hindered, though, because it has been difficult to directly compare results across studies and synthesize findings to understand the broader patterns of support.

>>>>

There has been considerable recent debate about the deepest split in the animal tree of life. This question has major implications for our understanding of the earliest events in animal evolution, including the origin of the nervous system. Some previous analyses found support for the first split being between comb jellies and all other animals (the Ctenophora-sister hypothesis), while others found support the first split being between sponges and all other animals (Porifera-sister). These studies differed in a many respects, including the genes and species selected, molecular evolution models, and analysis software. Here we systematically explore this question under consistent conditions by synthesizing data and results from XX previous phylogenomic analyses and performing new standardized analyses. Our reanalyses are consistent with previous studies that recover Porifera-sister with reduced outgroup (non-animal) sampling and complex site heterogeneous models, and Ctenophora-sister under other conditions. Though this had been interpreted by some to suggest Ctenophora-sister is an artifact of amino-acid composition bias, our analyses instead indicate that many site heterogeneous models do support Ctenophora-sister, and Porifera-sister is a result of unconstrained overfit models. Our analyses help reconcile variation across previous studies, and the datasets and analyses consolidated here will be a useful test-platform for phylogenetic methods development.

@@

Historically, there was little debate about the root of the animal tree of life. Porifera-sister (Fig. 1) was widely accepted though rarely tested. By contrast, there has long been uncertainty about the relationship of Ctenophora to other animals [@Wallberg:2004ws]. The first phylogenomic study to include ctenophores [@Dunn:2008ky] suggested a new hypothesis, now referred to as Ctenophora-sister, that ctenophores rather than sponges are our most distant animal relative. Since then many more studies have been published, with some analyses that find support for Ctenophora-sister, some for Porifera-sister, and some neither (Table 1). As the problem has become better characterized, it has become an interesting test-case for addressing difficult phylogenetic questions, which has attracted methodological interest well beyond investigators concerned with early animal evolution. These studies have taken a wide variety of approaches and results have varied across these different analyses. The extensive technical variation across these studies has been important to advancing our understanding of the sensitivity of these analyses, revealing for example that outgroup and model selection can have a large impact on result. But the larger technical variation has also made it difficult to synthesize these results to understand the details of this sensitivity because so many factors, such as which software tools are used, vary so widely across analyses. This variance is often confounded, for example when a particular model is only available in one software package. To advance this problem it is therefore critical to synthesize published results and explore variation in a standardized framework.
